---
title: "An integrated dataset of organohalogenated contaminants, trophic interactions and biological responses in the Seine estuary common sole nursery"
documentclass: article
link-citations: yes
author:
  - Florence MOUNIER:
      email: florence.mounier@inrae.fr
      institute: [INRAE]
      correspondence: true
  - Jérémy Lobry:
      institute: [INRAE]
  - Véronique Loizeau:
      institute: [LEMAR]
  - Florent Grasso:
      institute: [LEMAR]
  - Catherine Munschy: 
      institute: [Ifremer]
  - Yann Aminot:
      institute: [Ifremer]
  - Marine Gallien:
      institute: [LASEM]
  - Olivier Le Pape:
      institute: [DECOD_Rennes]
  - Jean-Baptiste Lecomte:
      institute: [DECOD_Rennes]
  - Etienne Rivot:
      institute: [DECOD_Rennes]
  - Léa Douchet:
      institute: [DECOD_Rennes]
  - Nathan Martin:
      institute: [DECOD_Rennes]
  - Juliette Champagnat:
      institute: [MARBEC]
  - Bastien Chouquet:
      institute: [CSLN]
  - Sylvain Duhamel:
      institute: [CSLN]
  - Jérôme Couteau:
      institute: [TOXEM]
  - Nicolas Bacq:
      institute: [GSA]
  - Hélène Budzinski:
      institute: [EPOC]
  - Pierre Labadie:
      institute: [EPOC]
institute:
  - INRAE: INRAE, UR EABX, 50 avenue de Verdun, 33612 Cestas Cedex, France
  - Ifremer: Ifremer, CCEM Contamination Chimique des Écosystèmes Marins, F-44000, Nantes, France
  - EPOC: Univ. Bordeaux, CNRS, UMR 5805 EPOC, 33405 Talence, France
  - LEMAR: PFOM/LARN, LEMAR UMR 6539, IFREMER, 29280 Plouzané, France
  - GSA: GIP Seine-Aval, 76176 Rouen, France
  - DECOD_Rennes: DECOD (Ecosystem Dynamics and Sustainability), Institut Agro, IFREMER, INRAE, 35042 Rennes, France
  - CSLN: Cellule de Suivi du Littoral Normand (CSLN), 53 rue de Prony, 76600 Le Havre, France
  - TOXEM: TOXEM, Montivilliers, France
  - LASEM: LASEM (Laboratoire d’analyses, de surveillance et d’expertise de la Marine), BCRM Toulon – BN TOULON/LASEM – BP 61, 83800 Toulon Cedex 9, France
  - MARBEC: MARBEC, Univ Montpellier, CNRS, Ifremer, IRD, Sète, France
output: 
  officedown::rdocx_document:
    reference_docx: template.docx
    tables:
      caption:
        style: "Table Caption"
    plots:
      caption:
        style: "Figure caption"
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: references.bib
number-sections: true
csl: "aquatic-living-resources.csl"
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: inline
---

<!--
%\VignetteIndexEntry{CHOPIN data descriptor}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 7,
  fig.height = 5,
  out.width = "95%"
)
library(chopin.data)
library(officer)
library(tidyverse)
library(flextable)
```

##### Abstract

Estuarine nurseries provide essential habitats for the early life stages of numerous marine fish species, but their ecological functioning can be affected by exposure to multiple chemical contaminants. Understanding contaminant accumulation and transfer within these habitats requires integrated observations of environmental contamination, benthic organisms, fish diet and biological responses. However, such data are often fragmented across environmental compartments and disciplinary approaches, limiting their reuse for trophic and ecotoxicological investigations.

Here, we present an integrated dataset assembled from the CHOPIN project (*Contaminants organoHalogénés histOriques et d’intérêt émergent : Présence et transfert vers la sole commune – Impact de la contamination sur la Nourricerie et conséquences sur la population*), conducted in the Seine estuary, France. The dataset characterizes contamination by polychlorinated biphenyls (PCBs), hexabromocyclododecane (HBCDD) isomers and per- and polyfluoroalkyl substances (PFAS) across surface sediments, benthic macrofauna and common sole (*Solea solea*).

The database combines contaminant concentrations and chemical reference information with sediment characteristics, benthic functional traits, carbon and nitrogen stable isotope signatures, juvenile sole stomach contents, fish morphometric measurements and biological biomarkers. Contaminant measurements in age-2 sole additionally distinguish liver, gonads and the remaining body compartment, enabling the investigation of tissue-specific accumulation and contaminant distribution.

The data are distributed as documented, directly usable R objects through the CHOPIN data package, together with preparation scripts and supporting documentation. This resource facilitates reproducible investigations of contaminant occurrence, bioaccumulation, trophic transfer, fish diet and biological responses in estuarine nursery habitats.


##### keywords

Seine estuary; common sole; benthic food web; PCB; PFAS; HBCDD; stable isotopes; bioaccumulation; ecotoxicology; R data package.

##### Article type

Data Descriptor

##### Target journal

Scientific Data

\newpage


## Background & Summary

Coastal and estuarine habitats support the juvenile stages of numerous marine fish species and contribute to the renewal of exploited populations. These nursery habitats are subject to multiple environmental pressures, including habitat modification and chemical contamination, which may affect the growth, survival and subsequent contribution of juvenile fish to adult populations.

The Seine estuary, located in northwestern France, provides nursery habitats for the common sole (*Solea solea*), a commercially important flatfish in the eastern English Channel. At the same time, the estuary receives contaminants originating from its catchment and local activities. Sediments can act as reservoirs of persistent organic contaminants, providing potential exposure pathways for benthic organisms and their predators.

The CHOPIN project was established to investigate the occurrence, accumulation and transfer of historical and emerging organohalogenated contaminants within the Seine estuary common sole nursery. Three contaminant families were considered: polychlorinated biphenyls (PCBs), hexabromocyclododecane (HBCDD) isomers and per- and polyfluoroalkyl substances (PFAS). The project also investigated biological responses in juvenile and subadult soles and explored the possible implications of nursery contamination for the population of common sole in the eastern English Channel.

The project integrated measurements from several environmental and biological compartments. Sediment sampling characterized the environmental contamination of nursery habitats, while contaminant analyses of benthic macrofauna provided information on potential dietary exposure. Measurements in common sole enabled comparisons among age classes and, for older individuals, among body compartments. Stable isotope signatures and stomach content analyses complemented chemical measurements by describing trophic relationships and prey consumption.

These observations provide several complementary perspectives on contaminant dynamics. Chemical concentrations can be used to characterize exposure and accumulation across compartments, while dietary information and stable isotope signatures support the investigation of trophic transfer. Tissue-specific measurements provide information on contaminant distribution within individual fish, and biological measurements allow associations between contaminant exposure and organism condition to be explored.

The present publication describes the data resources assembled from the CHOPIN project and distributed through a dedicated R package. Its purpose is to document their origin, structure, analytical scope, quality-control procedures and potential applications, rather than to test individual hypotheses regarding contaminant biomagnification or effects on fish populations.

The package brings together environmental, trophic, chemical and biological observations in a documented format, enabling researchers to access the different components of the dataset and combine them for further analyses.

## Methods

### Study area

The study was conducted in the lower Seine estuary, Normandy, France, which provides nursery habitats for common sole (Solea solea) during the juvenile phase of its life cycle. The CHOPIN project investigated the occurrence and transfer of organohalogenated contaminants across sediment, benthic macrofauna and common sole, with the aim of characterizing contaminant exposure within the estuarine nursery food web.

The study was conducted during several sampling campaigns in 2017 and 2018 in the lower Seine estuary and its adjacent marine environment in Normandy, France. The sampling design distinguished three principal estuarine sectors: the northern pit (*Fosse Nord*, FN), the southern pit (*Fosse Sud*, FS) and the estuary mouth (*Embouchure*, EM) (see Fig. \@ref(fig:fig-map)). The navigation channel was also included for selected biological samples. The northern and southern sectors are separated from the navigation channel by submerged embankments, whereas the Embouchure sector is located farther downstream, towards the adjacent marine environment. The CHOPIN sampling design also included additional location in the area offshore from Octeville for the study of older sole.

These sectors encompass spatial variation in hydrosedimentary conditions and benthic habitats. Sediment sampling focused on sandy to muddy substrates suitable for juvenile common sole and potentially favourable to the accumulation of sediment-associated contaminants.

Sediment and benthic macrofauna were sampled at eight principal stations, distributed among the three main sectors. Sampling stations were selected to represent fine sandy to muddy habitats used by juvenile sole and to provide biological material from benthic organisms known or expected to contribute to their diet.

Three main sampling campaigns were conducted: June 2017, October 2017 and October–November 2018. Sediment sampling covered all three campaigns, whereas benthic invertebrates were collected in spring and autumn 2017. Common sole were sampled during the same study period to obtain observations from successive juvenile age classes.


```{r fig-map, fig.cap="Map of the Seine estuary showing the sampling sectors and stations used for sediment sampling.", out.width="100%", fig.show='hold'}
knitr::include_graphics(
  system.file(
    "extdata",
    "map_seine.jpg",
    package = "chopin.data",
    mustWork = TRUE
  )
)
```


### Sediment sampling and characterization

Surface sediment was collected to characterize contaminant concentrations and selected physicochemical properties. Sampling locations were selected according to the local sediment characteristics, with particular emphasis on fine sediments corresponding to habitats used by juvenile common sole.

Surface sediment samples were collected using a grab sampler at eight stations distributed across three sectors of the estuary: Fosse Nord (FN), Fosse Sud (FS), and Embouchure (Emb), during three field campaigns (June and October 2017 and October 2018). Following collection, sediment samples were stored in stainless-steel containers in coolers until processing. Samples were homogenized using stainless-steel utensils and divided into aliquots according to the subsequent analytical procedures. Aliquots intended for contaminant and stable isotope analyses were freeze-dried and sieved through a 2-mm mesh. The processed material was subsequently stored in amber glass containers, protected from light, at room temperature. Sediment was homogenized again before analytical subsampling.

Sediment characteristics were measured to provide environmental information relevant to contaminant distribution and accumulation, including grain-size distribution, organic carbon and nitrogen contents, carbon-to-nitrogen ratio, porosity and water content, according to the availability of measurements.

Grain-size measurements were used to characterize the relative proportions of different sediment fractions, while organic carbon and nitrogen measurements provided information on the composition of sediment organic matter.

Individual contaminant concentrations are expressed relative to sediment dry weight. Organic carbon-normalized concentrations are also provided, together with summed concentrations and the relative contributions of individual compounds to their respective contaminant families.

The physicochemical characteristics, associated contaminant measurements, exact analytical definitions, measurement units and availability of each sediment variable are described in the package documentation related to the sed_contam dataset.


### Benthic macrofauna sampling and characterization

Benthic macrofauna was collected during the sediment sampling campaigns. Subtidal sampling was performed from fishing vessels using a Van Veen grab, whereas the intertidal sampling station was accessed on foot. Collected organisms were sorted by species and transferred to glass containers. Samples were maintained in coolers during transport to the laboratory. In the laboratory, organisms were rinsed with clean water, homogenized and freeze-dried. The resulting material was stored in amber glass containers, protected from light, at room temperature. Samples were homogenized again using stainless-steel utensils before analytical subsampling.

The sampling targeted species were selected according to their known occurrence in the Seine estuary, their potential contribution to juvenile sole diet and the quantity of biological material available for chemical analysis. Approximately 10 g of biological material per analytical sample was required to perform the planned measurements. The targeted assemblage comprised 16 benthic species belonging to three broad taxonomic groups: bivalves, polychaetes and crustaceans. Taxonomic identification and selected functional traits, including feeding mode and mobility, are provided through the benthos_info reference dataset. Benthic samples were characterized using dry matter content and lipid content, where available. Carbon and nitrogen stable isotope signatures were also measured for selected samples. These complementary data enable contaminant concentrations to be examined in relation to benthic taxonomy, trophic characteristics and potential prey identity. Contaminant concentrations and associated sample characteristics are distributed in benthos_contam.


### Common sole sampling and biological characterization

Common sole (Solea solea) were sampled to characterize changes in contaminant accumulation across juvenile age classes and successive birth cohorts within the Seine estuary nursery. Three sampling campaigns were conducted. Age-0 sole (G0) were collected in June 2017, shortly after recruitment to the estuary. The autumn 2017 campaign included both G0 and age-1 sole (G1), whereas G1 and age-2 sole (G2) were collected in autumn 2018.

Sample preparation and chemical analyses were adapted to fish age class. G0 individuals were pooled by sampling period and estuarine sector to obtain sufficient biological material for contaminant analysis. Mean individual length and body mass, together with their standard deviations, were recorded for each pool. G1 sole were analysed individually using whole-body samples. For G2 sole, sex was determined and each individual was dissected to separate the liver, gonads and remaining eviscerated body compartment. These compartments were analysed separately to characterize tissue-specific contaminant accumulation, including accumulation in reproductive organs, and to quantify contaminant distribution within individual fish.

In the resulting soles_contam dataset, G2 tissue-specific observations are distinguished by the sample_type variable, with the categories liver, gonads and rest. Measurements from the same individual are linked by a common sample identifier (sample_TAG), enabling comparisons of tissue-specific concentrations and reconstruction of contaminant burdens across the analysed body compartments.

Fish were measured and weighed before dissection and sample preparation. Biological samples were individually wrapped in pre-combusted aluminium foil and stored at −20 °C until processing. Whole-body samples or dissected tissues were subsequently homogenized and freeze-dried before chemical analysis. Freeze-dried material was stored in glass containers, protected from light, at room temperature. Morphometric characteristics, dry matter content and lipid content were recorded where available.

The dataset is complemented by stomach content and stable isotope measurements acquired through the CAPES project, providing additional information on the feeding ecology and trophic relationships of juvenile common sole.


### Target contaminants and chemical analyses

Three groups of organohalogenated contaminants were investigated: polychlorinated biphenyls (PCBs), hexabromocyclododecane (HBCDD) isomers and per- and polyfluoroalkyl substances (PFAS).

The analytical programme included 19 PCB congeners, the three α-, β- and γ-HBCDD isomers, and a panel of PFAS representing several chemical subclasses, including perfluoroalkyl carboxylic acids, perfluoroalkyl sulfonates, perfluorooctane sulfonamides and related derivatives, fluorotelomer sulfonates, disubstituted fluorotelomer phosphates and selected alternative substances.

The contam_info table documents compound identifiers, chemical families and subclasses, selected physicochemical properties and available analytical detection and quantification limits.

The set of compounds analysed differed among sediment and biological matrices. Consequently, comparisons among compartments should be restricted to chemicals measured in the corresponding samples.

Contaminant concentrations are provided for sediment and biological samples using the appropriate measurement basis. Dry-weight concentrations are available across the main analytical datasets. Additional organic carbon-normalized sediment concentrations and lipid-normalized biological concentrations are provided for selected compound families (PCBs and HBCDD isomers).


### Chemical analyses

#### PFAS extraction from biological samples

PFAS were extracted from freeze-dried and homogenized fish and benthic invertebrate samples following the analytical procedure described in the thesis of Marine Gallien (@gallienEtudeComparativeLecodynamique2021).

Approximately 250 mg of biological material was fortified with isotopically labelled internal standards before extraction. The internal standard mixture included labelled analogues of selected perfluoroalkyl carboxylic acids, perfluoroalkyl sulfonates, fluorotelomer sulfonates, sulfonamides and fluorotelomer phosphates.

Extraction was performed using 13 mL of methanol under microwave-assisted heating. The extraction programme consisted of heating to 70 °C over 5 min, maintaining this temperature for 5 min and cooling for 10 min.

The extracts were filtered through glass wool and concentrated under a nitrogen stream to approximately 2 mL.

Following dilution with 50 mL of ultrapure water, extracts were loaded onto weak anion-exchange solid-phase extraction cartridges (Strata X-AW).

After drying, the cartridges were connected to graphitized carbon cleanup cartridges (ENVI-Carb). Analytes were eluted using two successive 4-mL volumes of methanol containing 0.2% ammonium hydroxide.

Purified extracts were concentrated under nitrogen and transferred into polypropylene injection vials.

#### PFAS instrumental analysis

PFAS in biological samples were analysed by liquid chromatography coupled to tandem mass spectrometry.

The analytical system described in the thesis comprised an Agilent 1200 liquid chromatograph coupled to an Agilent 6490 triple-quadrupole mass spectrometer equipped with an Agilent Jet Stream electrospray ionization source.

Isotopically labelled internal standards were used for quantification and to account for analytical variability.

[The final manuscript should specify chromatographic separation conditions, ionization polarity, acquisition parameters and the quantification procedure where these details are required by the journal.]

#### PFAS extraction from sediment

PFAS extraction from sediment followed a procedure adapted to the sediment matrix.

Approximately 2 g of freeze-dried sediment was fortified with isotopically labelled internal standards and extracted using 10 mL of methanol under microwave-assisted heating.

As for biological samples, the microwave extraction programme reached 70 °C over 5 min, maintained this temperature for 5 min and included a 10-min cooling step.

The extracts were subsequently purified and concentrated before instrumental analysis.

PFAS in sediment were analysed using the same instrumental technique as that applied to biological samples.

The sediment procedure included matrix-specific quality-control samples and analytical limits.

#### PCB and HBCDD analyses

PCB and HBCDD analyses were performed using dedicated procedures for the corresponding chemical families and environmental matrices.

The HBCDD analytical programme distinguished the three principal diastereoisomers, α-, β- and γ-HBCDD, allowing isomer-specific concentration profiles to be characterized.

The thesis details the preparation and instrumental determination of HBCDD in sediment, while the CHOPIN project involved complementary analytical work conducted by the participating laboratories for different contaminants and biological matrices.

[To complete using the original laboratory protocols: Specify the extraction solvents, extraction and purification techniques, chromatographic instruments and quantification procedures used for PCB and HBCDD measurements in each biological matrix. The thesis does not provide all laboratory procedures for every compound–matrix combination included in the final package.]



[**To complete:** Specify the laboratory analytical procedures used for each chemical family and sample matrix, including extraction, purification, instrumental analysis, calibration, blank correction, internal standards, recoveries and treatment of values below the detection or quantification limits.]

#### Analytical quality assurance and quality control

Analytical quality assurance was implemented through the use of internal standards, procedural blanks, fortified control samples and reference materials.

For PFAS analyses of biological material, quality-control samples accompanied the analytical batches throughout the extraction and purification procedures.

These included fortified freeze-dried sole material, procedural blanks and the certified reference material SRM 1947 (Lake Michigan Fish Tissue, NIST).

The reference material was used to evaluate the analytical accuracy of selected PFAS measurements. Fortified biological samples were used to assess repeatability, and procedural blanks were used to characterize potential analytical contamination.

Measured quantities were corrected for the corresponding procedural blank where specified in the analytical protocol, then normalized to sample mass.

For sediment PFAS measurements, fortified Fontainebleau sand was used as an analytical control, in addition to procedural blanks.

The analytical quality-control programme also included the determination of compound-specific limits of detection and quantification.

The limits differed among contaminants and sample matrices. Accordingly, the interpretation of concentrations below analytical thresholds requires consideration of the relevant compound and matrix.

The contam_info reference table provides the detection and quantification limits retained in the published dataset.

[The final manuscript should include a concise table summarizing analytical quality-control results for each contaminant family and matrix, with the complete compound-specific limits supplied as supplementary information.]


### Stable isotope measurements

Carbon and nitrogen stable isotope signatures can be used to investigate variation in trophic resource use and relative trophic position among organisms.

Carbon and nitrogen stable isotope signatures were measured in selected benthic organisms and common sole from the CAPES and CHOPIN projects to characterize their trophic relationships and provide complementary information on the sources of organic matter supporting the benthic food web. Stable isotope measurements were also performed on sediment organic matter. 

The isotope dataset includes carbon and nitrogen isotope values (δ¹³C and δ¹⁵N), associated elemental composition measurements, taxonomic information, sampling metadata and selected biological characteristics.

The analytical results are distributed in the isotopes dataset, along with taxonomic information, sampling metadata and selected biological characteristics.

[**To complete:** Specify tissue preparation, lipid treatment if applicable, analytical instrumentation, isotope reference standards, measurement precision and any correction or normalization procedures. removal of inorganic carbon from sediment, analytical instrumentation, calibration standards and measurement precision.]


### Stomach content analysis

Stomach content data were obtained from the examination of juvenile common sole digestive tracts. Digestive tract contents were examined to identify prey consumed by juvenile common sole and characterize their relative contribution to the diet. Biometric and stomach content information was shared between the CAPES and CHOPIN projects to provide complementary information on the feeding ecology of juvenile sole and the identity of potential benthic prey.

The soles_stomach dataset contains individual fish characteristics, identified prey taxa, prey counts and observations associated with prey identification. Prey counts are recorded separately for the stomach and intestine, allowing diet composition to be examined using either compartment or the complete digestive tract, together with information on the degradation stage of the identified material. Taxonomic identification was retained at the lowest level supported by the observed material. Some prey could be identified to species, whereas others were assigned to genus or a higher taxonomic rank.

A supplementary taxonomic classification table, species_classif, provides the hierarchical classification of recorded prey taxa.

[**To complete:**: The dissection procedure, prey identification protocol, counting rules for fragmented organisms, and criteria used to assign degradation stagesmust be documented from the CAPES methodological material or the original stomach content analysis protocol. They should not be attributed to the chemical analytical methods described in Gallien's thesis.]


### Biological condition, histology and biomarkers

Biological measurements in common sole included morphometric characteristics, liver and gonad masses, hepatosomatic and gonadosomatic indices, histological observations and selected biomarkers. The hepatosomatic index (HSI) and gonadosomatic index (GSI) were calculated from liver and gonad wet masses relative to total fish wet mass, respectively. Histological observations of liver and gonadal tissue complemented these measurements.

Several biological biomarkers were measured to characterize different physiological and cellular responses. Acetylcholinesterase (AChE) activity was measured as a biomarker related to neurotoxicity, whereas ethoxyresorufin-O-deethylase (EROD) activity was measured as an indicator of biotransformation activity. Genotoxicity-related endpoints included the comet assay, expressed as Olive tail moment and percentage of DNA in the comet tail, as well as micronucleus measurements. Vitellogenin was measured as an indicator related to reproductive physiology. The dataset records AChE activity in µmol min⁻¹ mg⁻¹ and EROD activity in pmol min⁻¹ mg⁻¹. Other analytical units and histological classification codes are documented where available.

The resulting data are distributed in the soles_biomarkers dataset, together with fish morphometric characteristics, somatic indices and histological observations.

[**To complete:**: The precise biomarker and histological laboratory protocols, including tissue preparation, analytical replication, measurement units not yet specified and quality-control criteria, must be obtained from the CHOPIN project's relevant laboratory documentation.]


## Data compilation and preparation for the R package

The original CHOPIN analytical data and associated biological observations were compiled into a set of complementary datasets distributed through a dedicated R package, the R package **chopin.data**, which provides documented datasets and supporting reference tables. The data preparation workflow was designed to preserve sample identification, sampling metadata, analytical measurements and distinctions among environmental and biological matrices.

Separate datasets were created for sediment contamination, benthic contamination, common sole contamination, stable isotopes, stomach contents, biological biomarkers and taxonomic or chemical reference information. The datasets retain distinctions among individual fish, pooled biological samples and tissue-specific observations. These differences must be considered when combining tables or interpreting sample sizes. Variable names were harmonized across datasets to facilitate their use in R. Common identifiers include species, sample_TAG, year, season and zone, where applicable. The original chemical identifiers were retained across the contaminant datasets and linked to the reference table contam_info. Concentrations were expressed on a dry-weight basis, with additional lipid-normalized concentrations for selected biological contaminants and organic carbon-normalized concentrations for sediment. For age-2 sole, tissue-specific measurements were retained as separate observations linked by the individual sample identifier.

The source datasets and data processing scripts are provided in the package repository to document the transformations applied to the original data.

The package is organized into several complementary data domains. Each dataset is distributed as a directly accessible R object. Variable definitions, measurement units and examples of use are provided through the package documentation.


| Dataset            | Description                                                                                   |
| ------------------ | --------------------------------------------------------------------------------------------- |
| `sed_contam`       | Sediment characteristics and contaminant concentrations                                       |
| `benthos_contam`   | Benthic macrofauna characteristics and contaminant concentrations                             |
| `soles_contam`     | Common sole contaminant concentrations, including tissue-specific measurements in G2 fish     |
| `soles_stomach`    | Juvenile common sole digestive tract contents and prey observations                           |
| `isotopes`         | Carbon and nitrogen stable isotope signatures of benthos and sole                             |
| `soles_biomarkers` | Common sole morphometric measurements, somatic indices, histology and biomarkers              |
| `contam_info`      | Chemical identifiers, compound families, physicochemical properties and analytical parameters |
| `benthos_info`     | Benthic taxonomic and functional trait reference table                                        |
| `soles_info`       | Common sole taxonomic and functional trait reference table                                    |
| `species_classif`  | Taxonomic classification of organisms identified in sole digestive tract contents             |


Additional reference vectors are provided to facilitate the selection of contaminants and chemical subclasses in R analyses.

Several tables contain complementary observations referring to the same sampling locations, organisms or chemical compounds. The variables used to link these tables include sample identifiers, scientific names and chemical identifiers, as appropriate.

[To complete from the final package: Document the treatment of missing and censored values, conversion and normalization formulas, consistency checks, taxonomic harmonization, package version and archived data release.]

**Data and code availability:** [https://github.com/FlorenceMounier/chopin.data]


## Technical Validation

### Analytical quality control

[This section requires the analytical quality-control information associated with each compound family and biological matrix.]

The chemical reference table contains available limits of detection and quantification. These analytical thresholds should be considered when interpreting low contaminant concentrations and measurements below the quantification limit.

The procedures applied to non-detected compounds, censored values and missing measurements must be documented explicitly for each data table.


### Reproducibility of data preparation

[Describe the data preparation scripts supplied with the package, their input files and the transformations applied to the original CHOPIN database.]

The package documentation and preparation scripts should allow users to identify the origin of each distributed dataset and reproduce the transformation of the original analytical records into the published R objects.

## Usage Notes

The integrated dataset can support investigations of contaminant occurrence across environmental compartments, chemical accumulation in benthic organisms, trophic exposure of juvenile common sole and variation in contaminant distribution among fish tissues.

The combination of stomach content and stable isotope information enables comparisons between observed diet composition and broader trophic patterns. Together with contaminant concentrations, these data can support the estimation of sediment-to-biota accumulation factors, biomagnification factors and trophic magnification factors, subject to the assumptions of each approach.

The tissue-specific measurements from G2 sole provide opportunities to examine contaminant partitioning among liver, gonads and remaining body tissues. The biological measurements may also support exploratory analyses of associations between contaminant exposure, fish condition and biomarker responses.

Several limitations should be considered when reusing the data. Sampling coverage differs among compartments, years, sectors, species and fish age classes. Not all chemical compounds were analysed in every biological matrix. Pooled samples and tissue-specific measurements do not constitute independent observations of individual fish.

The dataset is therefore suited to analyses that explicitly account for sampling structure, measurement availability and the analytical characteristics of each contaminant family.

## References

[Insert the original CHOPIN project report and the methodological references relevant to sampling, analytical chemistry, stable isotope analysis, stomach content identification and biomarker measurements.]

## Acknowledgements

[To be completed with the participating institutions, laboratories, technical staff, funders and contributors to data collection and preparation.]

## Author Contributions

[To be completed according to the actual contributions to field sampling, laboratory analyses, database construction, software development, validation and manuscript preparation.]

## Competing Interests

The authors declare no competing interests.



